---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

# xdrxlsx

A small graphics device that writes R plots directly as OOXML DrawingML
shapes, for use with `openxlsx2::wb_add_drawing()`. No Cairo, FreeType,
fontconfig, or xml2 dependency.

## Install

```{r}
# install.packages("remotes")
remotes::install_github("JanMarvin/xdrxlsx")
```

## Usage

`xdr_xlsx()` opens the device and returns a file path. Draw with any R
plotting system, call `dev.off()`, then hand the file to
`openxlsx2::wb_add_drawing()`.

```{r}
library(xdrxlsx)
library(openxlsx2)

f <- xdr_xlsx(width = 6, height = 4)
plot(1:10, (1:10)^2, type = "b")
dev.off()

wb <- wb_workbook()$add_worksheet()$add_drawing(xml = f, dims = "A1")
```

### tinyplot

```{r}
library(tinyplot)

f <- xdr_xlsx(width = 6, height = 4)
plt(mpg ~ wt | factor(cyl), data = mtcars)
dev.off()

wb$add_worksheet()$add_drawing(xml = f, dims = "A1")
```

### ggplot2

```{r}
library(ggplot2)

f <- xdr_xlsx(width = 6, height = 4)
print(
  ggplot(mtcars, aes(wt, mpg, color = factor(cyl))) +
    geom_point() +
    theme_minimal()
)
dev.off()

wb$add_worksheet()$add_drawing(xml = f, dims = "A1")

if (interactive()) wb$open()
```

## Fitting a specific cell range

`wb_add_drawing(dims = "A1:G15")` repositions the drawing but does not
rescale it to the cell range's actual size. If you need it to match a
specific range, either anchor it to a single cell (`dims = "A1"`, which
preserves the exact rendered size), or rescale first:

```{r}
xdr_fit(f, width = 4.67, height = 3.13)
```

## Notes

- Text uses estimated font metrics, not real font data — layout that
  depends on precise text measurement (legend spacing, centered labels)
  may be slightly off, especially with unusual fonts.
- Fonts are referenced by name only, not embedded. The viewer needs the
  font installed, or it falls back silently.
- Gradient fills (`grid::linearGradient()`/`radialGradient()`) are
  supported; tiling patterns and clip paths/masks are not.
